---
title: "Test & Training MSE"
author: "Joon Bum Yang"
date: '2023-03-20'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaps)
library(tidyverse)
library(ISLR)
library(boot)
```


# Set up data set with p = 20 n = 1000 and quantative response vector of b_1 = b_2 = 2 and b_3 = b_4 = b_5 = 0.5
```{r}
p = 20
n = 1000
beta = c(c(2,2,2,0.5,0.5), rep(0,15))
set.seed(0)
X = matrix(rnorm(n*p),nrow = n,ncol = p)
xi = rnorm(n)
y = X %*% beta + xi

# Randomly split to 100 train and 900 test
indx = sample(1:n,100)
y.train = y[indx]
x.train = X[indx,]
y.test = y[-indx]
x.test = X[-indx,]
```

# Plotting Training MSE of each number of varaibles using best sub-set selection 
```{r}
reg.fit = regsubsets(x = x.train,y = y.train,nvmax = 20)
reg.summary = summary(reg.fit)
plot(reg.summary$rss/length(y.train),xlab="Number of Variables",ylab="MSE",type="l")

```
# Plotting Test set MSE of each number of varaibles using best sub-set selection  
```{r}
x.test = cbind(rep(1,900),x.test)
colnames(x.test) = names(coef(reg.fit,id=20))
test.mse <- rep(NA,20)
for(i in 1:20){
coefi <- coef(reg.fit,id=i)
pred <- x.test[,names(coefi)]%*%coefi
test.mse[i] <- mean((y.test-pred)^2)
}
plot(test.mse,xlab="Number of Variables",ylab="MSE",type="l")
```

# Compare the number of varaibles that produces mimimum testMSE
```{r}
opt.train = which.min(reg.summary$rss)
opt.test = which.min(test.mse)
```
The training MSE has the minimal value when the model size is 20, while the test MSE has the minimal
value when the model size is 5. The training error always decreases as the number of predictors increases.
Choosing the model based on the training MSE leads to overfitting. The selected model based on the test
MSE corresponds to the true model.

# Apply the same method with L2 Norm
```{r}
vals = rep(NA,20)
for(i in 1:20){
coefi <- coef(reg.fit,id=i)
betai = rep(0,21)
names(betai) = colnames(x.test)
betai[names(coefi)] = coefi
vals[i] = sqrt(sum((beta - betai[-1])^2))
}
plot(vals,xlab="Number of Variables",ylab="l2 norm",type="l")

```

As the number of variables increases, the ℓ2 norm of the estimation error first decreases, achieves its minimum
when the model size is 5, and then increases. It has the same trend as the test MSE plot

# Compare the result for forward Stepwise and backward stepwise selection
```{r}

#Forward Stepwise
reg.fit.forward = regsubsets(x = x.train,y = y.train,nvmax = 20,method="forward")
reg.summary.forward = summary(reg.fit.forward)

test.mse <- rep(NA,20)
for(i in 1:20){
coefi <- coef(reg.fit.forward,id=i)
pred <- x.test[,names(coefi)]%*%coefi
test.mse[i] <- mean((y.test-pred)^2)
}
which.min(reg.summary.forward$rss)
which.min(test.mse)

#Backward Stepwise selection
reg.fit.backward = regsubsets(x = x.train,y = y.train,nvmax = 20,method="backward")
reg.summary.backward = summary(reg.fit.backward)
test.mse <- rep(NA,20)
for(i in 1:20){
coefi <- coef(reg.fit.backward,id=i)
pred <- x.test[,names(coefi)]%*%coefi
test.mse[i] <- mean((y.test-pred)^2)
}
which.min(reg.summary.backward$rss)
which.min(test.mse)

```



